\vspace*{\fill}

\noindent
\hfill
\begin{minipage}{0.7\textwidth}
\raggedleft

I would like to thank my family and friends, and especially Javier Sempere Hernández, who I hope will soon be my husband, for their continuous support throughout this journey at the Universidad Politécnica de Madrid. I would also like to thank all my research colleagues and everyone who has collaborated with me. In particular, I am especially grateful to Miguel Marchamalo, Rubén Martínez, and Sergio Álvarez from the Topography Laboratory at the School of Civil Engineering for selecting me in my second year of the Bachelor’s degree, incorporating me as a researcher in their laboratory, and for their continued support since then. I would not have come this far without you, and I will always be deeply grateful for your support.

\end{minipage}

\newpage


In parallel with this thesis, the author has contributed to several research publications, conference presentations, and scientific collaborations.

\section*{Scientific Publications}

The following scientific publications have been published as part of this thesis, with the author of this thesis serving as first author:

\begin{itemize}

    \item \emph{A Methodology for the Automated Estimation of Seismic Behaviour Modifiers from Building Footprints}.\\
    Authors: Miguel Ureña Pliego, Javier Rodríguez Saiz, Beatriz González Rodrigo and Miguel Marchamalo Sacristán.\\
    Published in \textit{Advanced Modeling and Simulation in Engineering Sciences} (Springer), 2026.\\
    Q1 journal in Modelling and Simulation and Applied Mathematics.\\
    DOI: \url{https://doi.org/10.1186/s40323-026-00323-y}
    
    \item \emph{Transport-Related Surface Detection with Machine Learning: Analyzing Temporal Trends in Madrid and Vienna}.\\
    Authors: Miguel Ureña Pliego, Nianfang Shi, Takeru Shibayama, Ulrich Leth, and Miguel Marchamalo Sacristán.\\
    Published in \textit{Remote Sensing Applications: Society and Environment} (Elsevier), 2025.\\
    Q1 journal in Geography, Planning and Development and Computers in Earth Sciences.\\
    DOI: \url{https://doi.org/10.1016/j.rsase.2025.101503}
    
    \item \emph{Automatic Building Height Estimation: Machine Learning Models for Urban Image Analysis}.\\
    Authors: Miguel Ureña Pliego, Rubén Martínez Marín, Beatriz González Rodrigo, María Belén Benito Oterino, and Miguel Marchamalo Sacristán.\\
    Published in \textit{Applied Sciences} (MDPI), 2024.\\
    Q2 journal in Engineering (Miscellaneous).\\
    DOI: \url{https://doi.org/10.3390/app13085037}
    

\end{itemize}

\section*{Conference Contributions}

The following conference papers and presentations are related to this thesis:

\begin{itemize}
    \item \emph{Machine Learning Classification of Structural Systems Using Exclusively Geospatial and Remote Sensing Attributes: A Case Study in San José, Costa Rica}.\\
    Authors: Miguel Ureña Pliego, Javier Rodríguez Saiz, Beatriz Moya García, Sebastián Rodríguez Iturra, Beatriz González Rodrigo, and Miguel Marchamalo Sacristán.\\
    8th European Conference on Earthquake Engineering (ECCE 2026), Berlin, Germany, 14--18 September 2026.\\
    Accepted for presentation. To be presented by Miguel Ureña Pliego and Javier Rodríguez Saiz.
    
    \item \emph{Automated Characterization of Seismic Plan Irregularities and Geometric Behaviour Modifiers Based on Building Footprints: A Case Study in San José, Costa Rica}.\\
    Authors: Miguel Ureña Pliego, Javier Rodríguez Saiz, Gonzalo Núñez Álvarez, Miguel Marchamalo Sacristán, and Beatriz González Rodrigo.\\
    11th European Workshop on the Seismic Behaviour of Irregular and Complex Structures (EWICS 2026), Seville, Spain, 15--16 October 2026.\\
    Accepted for presentation. To be presented by Miguel Ureña Pliego and Javier Rodríguez Saiz.
    
    \item \emph{A Methodology for the Automated Estimation of Seismic Behaviour Modifiers from Building Footprints}.\\
    III EELISA Student Scientific Competition, Bucharest, Romania, 2025.\\
    Presented by Miguel Ureña Pliego. Awarded Second Prize.\\
    URL: \url{https://short.upm.es/sfs56https://short.upm.es/sfs56}
    
    \item 3rd IACM Digital Twins in Engineering Conference (DTE 2025) and 1st ECCOMAS Artificial Intelligence and Computational Methods in Applied Science Conference (AICOMAS 2025), Paris, France, 17--21 February 2025.\\
    No conference paper submission. Attended by Miguel Ureña Pliego.
    
    \item \emph{Automated Algorithm for Spatial Characterisation of Buildings within Urban Blocks}.\\
    Authors: Miguel Ureña Pliego, Javier Rodríguez Saiz, Beatriz González Rodrigo, and Miguel Marchamalo Sacristán.\\
    10th International Conference on Computational Methods in Structural Dynamics and Earthquake Engineering (COMPDYN 2025), Rhodes, Greece, 15--18 June 2025.\\
    Presented by Javier Rodríguez Saiz.\\
    DOI: \url{https://doi.org/10.7712/120125.12870.26844}
        
    \item \emph{Transport-Related Surface Detection with Machine Learning: Analyzing Temporal Trends in Madrid and Vienna}.\\
    II EELISA Student Scientific Competition, Budapest, Hungary, 2024.\\
    Presented by Miguel Ureña Pliego. Awarded First Prize.\\
    URL: \url{https://xplore.bme.hu/eelisa-student-scientific-competition-a-tradition-is-being-born/}
    
    \item \emph{Automatic Building Height Estimation: Machine Learning Models for Urban Image Analysis}.\\
    Authors: Miguel Ureña Pliego, Rubén Martínez Marín, Beatriz González Rodrigo, María Belén Benito Oterino, and Miguel Marchamalo Sacristán.\\
    18th World Conference on Earthquake Engineering (WCEE 2024), Milan, Italy, 2024.\\
    Presented by Beatriz González Rodrigo.\\
    URL: \url{https://proceedings-wcee.org/view.html?id=24498&conference=18WCEE}
    
    \item \emph{Conference Contribution Published in Seismological Research Letters}.\\
    Authors: Beatriz González Rodrigo et al.\\
    SSA Annual Meeting 2023, Puerto Rico, USA, 2023.\\
    Presented by Beatriz González Rodrigo.\\
    Published in \textit{Seismological Research Letters}, Vol. 94(2B), pp. 1029--1314.\\
    DOI: \url{https://doi.org/10.1785/0220230054}
    

\end{itemize}

\section*{Research Collaborations}

During the development of this thesis, the author collaborated with the following institutions and researchers:

\begin{itemize}

    \item \textit{Advanced Geomatics Group (AGA), Department of Land Morphology and Engineering, Escuela Técnica Superior de Ingenieros de Caminos, Canales y Puertos, Universidad Politécnica de Madrid, Spain}.\\
    Researchers: Miguel Marchamalo Sacristán, Rubén Martínez Marín, Javier Rodríguez Saiz, Rodrigo Hermosilla, Diego Macarrón, Francisco García, Sergio Álvarez, María Belén Benito Oterino and Gonzalo Núñez Álvarez. 
    \url{https://blogs.upm.es/aga/}

    \item \textit{Department of Forestry and Environmental Engineering and Management, Forest Engineering School, Universidad Politécnica de Madrid, Spain} \\ 
    Researcher: Beatriz González Rodrigo
    \url{https://web.montes.upm.es/}
    
    \item \textit{PIMM Laboratory, ENSAM Institute of Technology, Paris, France}.\\
    Researchers: Beatriz Moya García, Sebastián Rodríguez Iturra, and Francisco Chinesta Soria.
    \url{https://pimm.artsetmetiers.fr/en}
    
    \item \textit{Institute of Transportation, TU Wien, Austria}.\\
    Researchers: Nianfang Shi, Takeru Shibayama, and Ulrich Leth.
    \url{https://www.tuwien.at/en/cee/transport}
    
    \item \textit{City Science Group, MIT Media Lab, Massachusetts Institute of Technology, USA}.\\
    Researchers: Kent Larson, Luis Alonso, Naroa Coretti, Leticia Izquierdo and Yasushi Sakai.
    \url{https://www.media.mit.edu/groups/city-science/overview/}
    
    \item \textit{Instituto Tecnológico de Santo Domingo (INTEC), Santo Domingo, Dominican Republic}.\\
    Researcher: Claudia Germoso.
    \url{https://www.intec.edu.do/en/}
    
    \item \textit{Universidad de Costa Rica (UCR), San José, Costa Rica}.
    Researcher: Diego Hidalgo.
    \url{https://www.ucr.ac.cr/}
    
    \item \textit{Universidad de San Carlos de Guatemala (USAC), Ciudad de Guatemala, Guatemala}.
    Researcher: Omar Flores.
    \url{https://www3.usac.edu.gt/}
    
    \item \textit{Detektia}, Spain. \\
    Researchers: Candela Sancho, Alfredo Landa.
    \url{https://detektia.com/en/}
    
    \item \textit{Geolyder}, Spain. \\
    Researchers: Orlando Hernández.
    \url{https://geolyder.com/}
        
\end{itemize}

\section*{Funding and Research Projects}

This research was supported by the following projects and funding programmes:

\begin{itemize}

    \item European Union Project 9063-03 \emph{Adelante 2}: \emph{Sustainable and Resilient Construction in Central America and the Caribbean in the Face of Seismic Hazard: Regional Cooperation Based on the Experience of Costa Rica}.
    
    \item \emph{SIAGUA Project} (PID2021-128123OB-C22), funded by MCIN/AEI/10.13039/501100011033 and the European Regional Development Fund (ERDF), ``A Way of Making Europe''.
    
    \item \emph{RISK-CARIBERIA Project} (PID2024-155792OB-I00), funded by the Spanish State Research Agency (AEI) under the 2025 Knowledge Generation Call of the Spanish State Plan for Scientific, Technical and Innovation Research.
    
    \item Industrial Doctorates Programme of the Community of Madrid (Grant IND2023/TIC-28743).
    
    \item Erasmus+ Programme.

    \item UPM scholarship to fund short research stays at MIT.

    \item Detektia.
    
\end{itemize}
